---
title: "ABA Reports 2018"
author: "Tanner Gildea"
date: "5/21/2019"
output: html_document
---

# ABA Reports 2018

Each year, the American Bar Association requires law schools to release standardized information regarding their class profiles and admissions practices. These annual ABA reports shed useful insight into top American law schools, including information on acceptance rates, average LSAT and GPA scores, average scholarship and grant aid amounts, and the diversity of particular law school classes.

As a potential future law school applicant, I wanted to analyze some of this data for myself. I accessed aggregate ABA reports from 2018 - below are some visualization of the results of my analysis.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(ggplot2)
library(readxl)
library(janitor)
library(ggthemes)
library(scales)

# Load in the two excel spreadsheets from the ABA.

first_year <- read_excel("2018_First_Year_Class.xlsx") %>% 
  clean_names() %>% 
  mutate(school_name = str_to_title(school_name)) 

grants <- read_excel("2018_Grants_and_Scholarships.xlsx") %>% 
  clean_names() %>% 
  mutate(school_name = str_to_title(school_name))
```

# Most Competitive Law Schools


```{r acceptance rate}
###############
# ACCEPTANCE RATE 
###############

first_year %>% 
  arrange(acceptance_rate) %>% 
  slice(1:20) %>% 
  
  # start visualization
  
  ggplot(aes(x= reorder(school_name, acceptance_rate), 
             y = acceptance_rate,
             fill = "Blue")) +
  
  geom_col() + 
  
  geom_text(aes(label= u_s_news_and_world_ranking)) +
  
  # theme changes
  
  theme_economist() +
  scale_fill_economist() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  
  
  labs(x = NULL,
       y = "Acceptance Rate",
       title = "Most Competitive Law Schools") +
  
  guides(fill = FALSE)
```

# Law Schools with the Highest Average GPA's

```{r GPA}
###############
# GPA's
###############

first_year %>% 
  arrange(desc(x50th_percentile_ugpa_all)) %>% 
  slice(1:20) %>% 
  
  # start visualization
  
  ggplot(aes(x= reorder(school_name, x50th_percentile_ugpa_all), 
             y = x50th_percentile_ugpa_all,
             fill = "yellow")) +
  
  geom_col() +
  
  geom_text(aes(label= u_s_news_and_world_ranking)) +
  
  # theme changes
  
  theme_economist() +
  scale_fill_economist() +
  coord_cartesian(ylim = c(3.5,4)) +
  
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +

  
  labs(x = NULL,
       y = "Median GPA",
       title = "Median Accepted Undergraduate GPA's") +
  
  guides(fill = FALSE) 
  
  
```

# Law Schools with the Highest Average LSAT scores

```{r LSAT scores}
###############
# LSAT scores
###############

first_year %>% 
  arrange(desc(x50th_percentile_lsat_all)) %>% 
  slice(1:20) %>% 
  
  # start visualization
  
  ggplot(aes(x= reorder(school_name, x50th_percentile_lsat_all), 
             y = x50th_percentile_lsat_all,
             fill = "Blue")) +
  
  geom_col() + 
  
  geom_text(aes(label= u_s_news_and_world_ranking)) +
  
  # theme changes
  
  theme_economist() +
  scale_fill_economist() +
  coord_cartesian(ylim = c(160,175)) +
  
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  
    labs(x = NULL,
       y = "Median LSAT Score",
       title = "Median Accepted LSAT Score") +
  
  guides(fill = FALSE)
```

# U.S. News and World Rankings 


```{r ranking }
###############
# US News and World Ranking
###############

first_year %>% 
  arrange(u_s_news_and_world_ranking) %>% 
  slice(1:20) %>% 
  
  # start visualization
  
  ggplot(aes(x= reorder(school_name, u_s_news_and_world_ranking), 
             y = u_s_news_and_world_ranking,
             fill = "Blue")) +
  
  geom_col() + 
  
  # theme changes
  
  theme_economist() +
  scale_fill_economist() +
  
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  
    labs(x = NULL,
       y = "Ranking",
       title = "Top Law Schools Per 2020 U.S. News Rankings") +
  
  guides(fill = FALSE) 
```