---
title: "ABA Reports 2018"
author: "Tanner Gildea"
date: "5/21/2019"
output: html_document
---

# ABA Reports 2018

Each year, the American Bar Association requires law schools to release standardized information regarding their class profiles and admissions practices. These annual ABA reports shed useful insight into top American law schools, including information on acceptance rates, average LSAT and GPA scores, average scholarship and grant aid amounts, and the diversity of particular law school classes.

As a potential future law school applicant, I wanted to analyze some of this data for myself. I accessed aggregate ABA reports from 2018 - below are some visualization of the results of my analysis.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(ggplot2)
library(readxl)
library(janitor)
library(ggthemes)
library(scales)

# Load in the four excel spreadsheets from the ABA.

# Selectivity, GPA, LSAT data

first_year <- read_excel("2018_First_Year_Class.xlsx") %>% 
  clean_names() %>% 
  mutate(school_name = str_to_title(school_name)) 


# Scholarship and Grant data

grants <- read_excel("2018_Grants_and_Scholarships.xlsx") %>% 
  clean_names() %>% 
  mutate(school_name = str_to_title(school_name)) %>% 
  
  # create percent of half or more grant aid
  
  mutate(half_plus_aid = half_to_full_tuition_total_percent + 
           full_tuition_total_percent + 
           more_than_full_tuition_total_percent)


# Ethnicity data

ethnicity <- read_excel("2018_JD_Enrollment_and_Ethnicity-2.xlsx") %>% 
  clean_names()  %>% 
  rename(school_name = school_list) %>%
  mutate(school_name = str_to_title(school_name)) %>% 
  mutate(percent_minority = (minority_degrees_awd / total_degrees_awarded)*100) %>% 
  mutate(percent_white = (white_jd_degrees_awd / total_degrees_awarded)*100) %>% 
  mutate(percent_international = (nonresident_alien_jd_degrees_awd / total_degrees_awarded)*100) %>%
  
  # great rounding feature!
  
  mutate_if(is.numeric, round, 2)


# Employment and Post-grad Data

employment <- read_excel("Employment Summary_2019.xlsx") %>% 
  clean_names() %>% 
  mutate(school_name = str_to_title(school_name)) %>% 
  mutate(first_largest_employment = str_to_title(first_largest_employment)) %>% 
  mutate(second_largest_employment = str_to_title(second_largest_employment)) %>% 
  mutate(third_largest_employment = str_to_title(third_largest_employment)) 


# Join all datasets to the master "all" set

all <- left_join(first_year, grants, by = "school_name") 
  
all <- left_join(all, ethnicity, by = "school_name")

all <- left_join(all, employment, by = "school_name")


# Write all datasets to rds files

write_rds(all, "ABA_Reports/all")

write_rds(ethnicity, "ABA_Reports/ethnicity")
write_rds(grants, "ABA_Reports/grants")
write_rds(first_year, "ABA_Reports/first_year")
write_rds(employment, "ABA_Reports/employment")

```

# Most Competitive Law Schools


```{r acceptance rate}
###############
# ACCEPTANCE RATE 
###############

first_year %>% 
  arrange(acceptance_rate) %>% 
  slice(1:20) %>% 
  
  # start visualization
  
  ggplot(aes(x= reorder(school_name, acceptance_rate), 
             y = acceptance_rate,
             fill = "Blue")) +
  
  geom_col() + 
  
  geom_text(aes(label= u_s_news_and_world_ranking)) +
  
  # theme changes
  
  theme_economist() +
  scale_fill_economist() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  
  
  labs(x = NULL,
       y = "Acceptance Rate",
       title = "Most Competitive Law Schools") +
  
  guides(fill = FALSE)
```

# Law Schools with the Highest Average GPA's

```{r GPA}
###############
# GPA's
###############

first_year %>% 
  arrange(desc(x50th_percentile_ugpa_all)) %>% 
  slice(1:20) %>% 
  
  # start visualization
  
  ggplot(aes(x= reorder(school_name, x50th_percentile_ugpa_all), 
             y = x50th_percentile_ugpa_all,
             fill = "yellow")) +
  
  geom_col() +
  
  geom_text(aes(label= u_s_news_and_world_ranking)) +
  
  # theme changes
  
  theme_economist() +
  scale_fill_economist() +
  coord_cartesian(ylim = c(3.7,4)) +
  
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +

  scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
  
  labs(x = NULL,
       y = "Median GPA",
       title = "Median Accepted Undergraduate GPA's") +
  
  guides(fill = FALSE) 
  
  
```

# Law Schools with the Highest Average LSAT scores

```{r LSAT scores}
###############
# LSAT scores
###############

first_year %>% 
  arrange(desc(x50th_percentile_lsat_all)) %>% 
  slice(1:20) %>% 
  
  # start visualization
  
  ggplot(aes(x= reorder(school_name, x50th_percentile_lsat_all), 
             y = x50th_percentile_lsat_all,
             fill = "Blue")) +
  
  geom_col() + 
  
  geom_text(aes(label= u_s_news_and_world_ranking)) +
  
  # theme changes
  
  theme_economist() +
  scale_fill_economist() +
  
  coord_cartesian(ylim = c(164,175)) +
  
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) +
  
    labs(x = NULL,
       y = "Median LSAT Score",
       title = "Median Accepted LSAT Score") +
  
  guides(fill = FALSE)
```

# U.S. News and World Rankings 


```{r ranking }
###############
# US News and World Ranking
###############

first_year %>% 
  arrange(u_s_news_and_world_ranking) %>% 
  slice(1:20) %>% 
  
  # start visualization
  
  ggplot(aes(x= reorder(school_name, u_s_news_and_world_ranking), 
             y = u_s_news_and_world_ranking,
             fill = "Blue")) +
  
  geom_col() + 
  
  # theme changes
  
  theme_economist() +
  scale_fill_economist() +
  
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  
    labs(x = NULL,
       y = "Ranking",
       title = "Top Law Schools Per 2020 U.S. News Rankings") +
  
  guides(fill = FALSE) 
```

# Median Grant Award at Top 20 Schools

```{r median_grant }
###############
# Financial Assistance by Selectivity
###############

all %>% 
  arrange(u_s_news_and_world_ranking) %>% 
  slice(1:20) %>% 
  
  # start visualization
  
  ggplot(aes(x= reorder(school_name, ft_50th_percentile_grant_amount), 
             y = ft_50th_percentile_grant_amount,
             fill = "Blue")) +
  
  geom_col() + 
  
  geom_text(aes(label= u_s_news_and_world_ranking)) +
  
  # theme changes
  
  theme_economist() +
  scale_fill_economist() +
  
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  
  coord_cartesian(ylim = c(10000,45000)) +
  
    scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
  
    labs(x = NULL,
       y = "Dollars",
       title = "Median Grant Amount") +
  
  guides(fill = FALSE) 

```

# Percent of Students Receiving More than Half Tuition 

```{r percent_aid }
###############
# Financial Assistance by Selectivity
###############

all %>% 
  arrange(u_s_news_and_world_ranking) %>% 
  slice(1:20) %>% 
  
  # start visualization
  
  ggplot(aes(x= reorder(school_name, half_to_full_tuition_total_percent), 
             y = half_to_full_tuition_total_percent,
             fill = "Blue")) +
  
  geom_col() + 
  
  geom_text(aes(label= u_s_news_and_world_ranking)) +
  
  # theme changes
  
  theme_economist() +
  scale_fill_economist() +
  
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  
  coord_cartesian(ylim = c(0,50)) +
  
    labs(x = NULL,
       y = "Percent",
       title = "Percent Receiving Half to Full Tuition in Grant Aid") +
  
  guides(fill = FALSE) 

```

# Percent of Students Receiving Full Tuition 

```{r full_percent_aid }
###############
# Financial Assistance by Selectivity
###############

all %>% 
  arrange(u_s_news_and_world_ranking) %>% 
  slice(1:20) %>% 
  
  # start visualization
  
  ggplot(aes(x= reorder(school_name, full_tuition_total_percent), 
             y = full_tuition_total_percent,
             fill = "Blue")) +
  
  geom_col() + 
  
  geom_text(aes(label= u_s_news_and_world_ranking)) +
  
  # theme changes
  
  theme_economist() +
  scale_fill_economist() +
  
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  
  coord_cartesian(ylim = c(0,20)) +
  
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
  
    labs(x = NULL,
       y = "Percent",
       title = "Percent Receiving Full Tuition in Grant Aid") +
  
  guides(fill = FALSE) 

```

# Percent of Students Receiving More than Full Tuition 

```{r full_plus_percent_aid }
###############
# Financial Assistance by Selectivity: More than Full Tuition
###############

all %>% 
  arrange(u_s_news_and_world_ranking) %>% 
  slice(1:20) %>% 
  
  # start visualization
  
  ggplot(aes(x= reorder(school_name, more_than_full_tuition_total_percent), 
             y = more_than_full_tuition_total_percent,
             fill = "Blue")) +
  
  geom_col() + 
  
  geom_text(aes(label= u_s_news_and_world_ranking)) +
  
  # theme changes
  
  theme_economist() +
  scale_fill_economist() +
  
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  
  coord_cartesian(ylim = c(0,10)) +
  
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) +
  
    labs(x = NULL,
       y = "Percent",
       title = "Percent Receiving More than Full Tuition in Grant Aid") +
  
  guides(fill = FALSE) 

```

```{r half_plus_aid }
###############
# Total percent receiving half or more aid
###############

all %>% 
  arrange(u_s_news_and_world_ranking) %>% 
  slice(1:20) %>% 
  
  # start visualization
  
  ggplot(aes(x= reorder(school_name, half_plus_aid), 
             y = half_plus_aid,
             fill = "Blue")) +
  
  geom_col() + 
  
  geom_text(aes(label= u_s_news_and_world_ranking)) +
  
  # theme changes
  
  theme_economist() +
  scale_fill_economist() +
  
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  
  coord_cartesian(ylim = c(0,70)) +
  
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
  
    labs(x = NULL,
       y = "Percent",
       title = "Percent Receiving More than Half Tuition in Grant Aid") +
  
  guides(fill = FALSE) 

```


# Percent of Students Receiving Any Amount of Aid

```{r percent_any_aid }
###############
# Financial Assistance by Selectivity: Any Aid
###############

all %>% 
  arrange(u_s_news_and_world_ranking) %>% 
  slice(1:20) %>% 
  
  # start visualization
  
  ggplot(aes(x= reorder(school_name, total_number_receiving_grants_total_percent), 
             y = total_number_receiving_grants_total_percent,
             fill = "Blue")) +
  
  geom_col() + 
  
  geom_text(aes(label= u_s_news_and_world_ranking)) +
  
  # theme changes
  
  theme_economist() +
  scale_fill_economist() +
  
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  
  coord_cartesian(ylim = c(30,100)) +
  
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +

  
    labs(x = NULL,
       y = "Percent",
       title = "Percent of Students Receiving Any Grant Aid") +
  
  guides(fill = FALSE) 

```

# Diversity profiles of top 20 schools
```{r ethnicity}
###############
# Ethnicity Top 20
###############





```

